\textls[-0]{\thetitle}

### 2023-03-18 metatheory2.tex
pass over para on universe levels

\subsection{Review MLTT in Debruijn Indices, Explicit Subst, and Explicit Universe Level}

\TT can be roughly considered as MLTT with linkages so we still start with vanilla MLTT first.

\input{all-mltt-typing}

@@

\noindentparagraph{Explicit Universe Level.} Our treatment of universe level mainly follows that of \citet{altkap2016, kaposi2019gluing}. Universe level is mainly to get around the size issue---we cannot refer to(and thus quantifiy over) a set of all set (or a universe of all type). So everytime we want to quantify over a universe of types, we have to specify ``how large'' that universe is, that is basically what universe level is measuring.

Now we will need to some attach syntax pieces with natural number indicating universe level. The explicit universe level $i$ now will attach to each context and type judgement $\goodCtx{\Gm}{i}$ and $\goodType{\Gm}{T}{i}$ basically indicating the context $\Gm$ and $T$ inhabits at what universe. Apparently, universe level will also be attached to $\cU$. For two universe level $i,j$ (as natural number), we use $i \lcup j$ to denotes $\max(i,j)$.

Since we work in an intrinsic-typed setting, some required premises can be omitted without ambiguity. For omitted context judgement $\goodCtx{\Gm}{k}$, the universe level is always $k$, hinted at the figure of judgement specification. Similar for type judgement $\goodType{\Gm}{T}{j}$.

>>>>

\subsection{MLTT with Explicit Substitutions and Universe Levels}

We review the base MLTT fragment of \TT first.
\input{all-mltt-typing}

@@

\noindentparagraph{Universe levels.}

Universe levels address the size issue---it is unsound to have a set of all sets
(or a universe of all types).
The level of a universe specifies ``how large'' that universe is.
%that is basically what universe level is measuring.

%The treatment of universe levels follows \citet{altkap2016, kaposi2019gluing}.
The judgment form $\goodType{\Gm}{T}{i}$ indicates that (the code of) the type $T$
inhabits universe~$\cU_i$.
Similarly, $\goodCtx{\Gm}{i}$ indicates that (the codes of) the types in $\Gm$
inhabit universe~$\cU_i$.
The notation $i \lcup j$ denotes $\max(i,j)$.

%Since we work in an intrinsic-typed setting, some required premises can be omitted without ambiguity. For omitted context judgement $\goodCtx{\Gm}{k}$, the universe level is always $k$, hinted at the figure of judgement specification. Similar for type judgement $\goodType{\Gm}{T}{j}$.
